---
title: "Titanic"
author: "John Russell"
date: "1/7/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

When examining the improbable tragedy of the sinking of the Titanic in 1912, it is a reasonable assumption to believe that there was a predictable pattern to who on board survived and who died. It is my belief that women and younger passengers had a greater chance of surviving due to those groups likely taking priority when filling the lifeboats. A thorough analysis will be conducted to test this hypothesis.

```{r, include = F}
library(tidyverse)
train_data <- read_csv('train.csv')
attach(train_data)
```

## Analysis
```{r}
glimpse(train_data)
```

A quick look at the data shows that there are 891 passengers as well as 12 different variables, including whether or not the passenger survived.

```{r}
train_data <- train_data %>%
  select(-c(Cabin, Ticket))
```
 We quickly notice that the Cabin number of passengers in rarely present, so we will omit this variable from the analysis as it does not provide any useful information.
 
```{r}
ggplot(train_data) +
  geom_bar(aes(Survived, fill = Sex, position = "fill"))
```
 
A quick visualization of the data shows that the majority of the passengers that survived are, in fact, female. 

To properly examine age, it will be grouped as infant (0-5), child (6-12), teenager (13-19),  young adult (20-29), adult (30-49), and elderly (50+). Passengers with unknown age will be ommitted.

```{r}
train_data <- train_data %>%
  filter(!is.na(Age)) %>%
  mutate(Age_Group = cut(Age, breaks = c(0, 5, 12, 19, 29, 49, 80), labels = c("infant", "child", "teenager", "young adult", "adult", "elderly")))
```
```{r}
ggplot(train_data) +
  geom_bar(aes(Age_Group, fill = Age_Group))
```

